Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-10-15 manuscript/manuscript.tex
futzing for resubmit

\begin{document}
\maketitle

%%%%%%%%%% START AFTER TITLE PARTIAL %%%%%%%%%%%%%

@@

Imagine you are a graduate student, and you run a replication of a study that you are interested in building upon in your research.
The replication fails.
Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a \(p\)-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to build upon?

In psychology, large scale replication projects have found that around half of studies successfully replicate.
Across 100 studies with positive results, the reproducibility project in psychology (RP:P) replicated 36\%-47\% depending on the metric for replication success (\citeproc{ref-opensciencecollaboration}{Open Science Collaboration, 2015}).
With large multi-site samples, Many Labs 1 replicated 11/13 effects (84.6\%), Many Labs 2 replicated 14/28 effects (50\%), and Many Labs 3 replicated 3/10 effects (30\%) (\citeproc{ref-ebersole2016}{Ebersole et al., 2016}; \citeproc{ref-klein2014}{Klein et al., 2014}; \citeproc{ref-klein2018}{Klein et al., 2018}).
Camerer et al. (\citeproc{ref-camerer2018}{2018}) examined 21 behavioral social science studies and successfully replicated 12-14 of them (57\%-67\%) depending on the metric used.
Boyce et al. (\citeproc{ref-boyce2023}{2023}) reported an average subjective replication score of 49\% for 176 replications primary in psychology.

Psychology is not the only discipline where many studies do not replicate.
Large scale replication projects in other disciplines have found 39/97 of studies with positive effects (40\%) replicating in cancer biology (\citeproc{ref-errington2021}{Errington et al., 2021}), 11/18 studies (61\%) replicating in economics (\citeproc{ref-camerer2016}{Camerer et al., 2016}), and 31/40 studies (78\%) replicating in experimental philosophy (\citeproc{ref-cova2021}{Cova et al., 2021}).
Across these disciplines, and especially in psychology, large-scale replications indicate a substantial chance of scientists encountering failed replications if they try to replicate published findings.

Replications can fail for many reasons, having to do with the original study or the replication.
Starting with the original study, one potential reason for replication failure is that the original result is fragile in some way.
It could be a statistically unlikely result that was achieved through chance or \(p\)-hacking, or it could be sensitive to the exact conditions and time it was run under (``hidden moderators'').
Another possibility is that the original reported effect size might be inflated due to some combination of heterogeneity, \(p\)-hacking, low power, and publication bias.
A consequence of inflated effect sizes is that replication studies might select a sample size with appropriate statistical power for the reported effect size and then have too small a sample to reliably detect the (true) smaller effect.
Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Some other reasons for replication failure are primarily due to the replication study.
A replication could observe different results from the original due to methodological differences.
These differences in design could be intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that were not expected to be critical.
A difference in the details of recruitment (e.g., online versus in-person) might mean that particular details of a study implementation would no longer be appropriate, or the data quality controls and attention checks might be inadequate.
If there are differences in time, place, or subject population between the two studies, corresponding changes to the materials, instructions, or procedure could be needed to provide an adaptation of the experimental paradigm to the new context.
Such adaptation creates yet more challenges in interpretation: too many changes could cause differing results, but so could not enough changes.
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

For the goals of large-scale replication projects, it does not matter why any particular study failed to replicate, because the aim is to estimate the proportion of replication effects in a literature.
In contrast, however, when individual scientists attempt to replicate an effect with the goal of building on it, they do care about why an individual replication failed.
In an uncertain literature, scientists might not want to commit resources to studying extensions of an effect without first checking that they can measure that effect themselves.
For those individual scientists, it is much more important whether the replication failure is due to a fixable issue with the replication or an intrinsic feature of the original result.

Re-replication can serve as a way to test explanations for replication failure.
While it is possible to speculate after the fact about potential causes of a single replication failure, in most circumstances these causes are hard to discern (absent clear evidence of errors).
In contrast, a second replication provides a chance to test explanations by making modifications -- as well as giving another chance for researchers to measure an effect of interest.
But what are the chances that a re-replication succeeds?

Relatively few studies have estimated re-replication success.
One estimate comes from Ebersole et al. (\citeproc{ref-ebersole2020}{2020}).
After the original RP:P study reported of a replication rate around 40\% (\citeproc{ref-opensciencecollaboration}{Open Science Collaboration, 2015}), Gilbert et al. (\citeproc{ref-gilbert2016}{2016}) raised concerns that methodological differences might explain the low replication rate.
In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication.
Thus, Ebersole et al. (\citeproc{ref-ebersole2020}{2020}) re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts.
This study thus tested whether larger sample sizes and/or ``better'' methodologies would ``rescue'' the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished (\citeproc{ref-ebersole2020}{Ebersole et al., 2020}).
The re-replication success rate was thus 2/9 (22\%) for studies with non-significant first replications, under favorable, high-resource circumstances.

While Ebersole et al. (\citeproc{ref-ebersole2020}{2020}) looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts.
On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure.
Thus, it is an open question what the re-replication rate is for failed replications by early career researchers.
Further, estimating this rate could inform researchers' decisions on what to do next after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate-level class on experimental methods in psychology were given the option to ``rescue'' an experiment that a student in a prior year had previously failed to replicate as their class project (first replications reported in \citeproc{ref-boyce2023}{Boyce et al., 2023}).
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project.
In the present paper, we report the results of 17 re-replication rescue projects.
We find that a minority of projects (5/17, 29\%) successfully recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students have conducted replication projects (\citeproc{ref-boyce2023}{Boyce et al., 2023}; \citeproc{ref-hawkins2018}{Hawkins et al., 2018}).
In Fall 2023, students were offered the option to complete a ``rescue'' project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for reproducibility, and 3) reviewing and approving the final manuscript.
We report on the result of 17 rescue projects that opted to be part of the paper and completed data collection.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), individual project data and analyses for rescue projects, and the analytic code for this paper are all available at \url{https://osf.io/cyk5w/}.

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is pre-registered at \url{https://osf.io/5qz7v}, individual pre-registrations are linked from \url{https://osf.io/vkbfw}.
We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance and the variance of the comparison (replication or rescue) study (\citeproc{ref-patil2016}{Patil et al., 2016}).
Thus we cannot show a single prediction interval for the original study, but would have to show a prediction interval between each pair of studies.
We thought this analysis would not offer additional clarity and hence we did not perform it.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (\url{https://osf.io/dfwk2}).
In general, we encouraged students to add manipulation checks as appropriate, and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of Paxton et al. (\citeproc{ref-paxton2012}{2012}) switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2 which is less overused but attempts to measure the same construct (\citeproc{ref-thomson2016}{Thomson \& Oppenheimer, 2016}).
The rescue of Jara-Ettinger et al. (\citeproc{ref-jara-ettinger2022}{2022}) discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
Tarampi et al. (\citeproc{ref-tarampi2016}{2016}) had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants click a drop down menu to select left or right for each turn.
The rescue removed the unwieldy drop down menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

With one exception, samples were collected on Prolific (the rescue of \citeproc{ref-yeshurun2003}{Yeshurun \& Levy, 2003} ran in-person on the Stanford student subject pool).
We tried to power studies adequately (with a target of 2.5x original following \citeproc{ref-simonsohn2015}{Simonsohn, 2015}), but due to cost constraints, not all studies were powered at this level.
The rescues had on average 1.48 times the original sample post-exclusion (median: 1.07, IQR: 0.94 - 2.4, minimum: 0.48, maximum: 2.96, see Table \ref{tab:tab-size} for all sample sizes).
Across the 16 Prolific studies, we spent \$5471, for an average of \$342 per project.

@@

As a complement to the subjective rating of overall success, we statistically compared one key measure of interest for each study, following Boyce et al. (\citeproc{ref-boyce2023}{2023}).
In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, first replication, and rescue.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

@@

We recorded the same set of potential correlates that were used in Boyce et al. (\citeproc{ref-boyce2023}{2023}) for original, first replication, and rescue (these were already rated for original and first replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between-subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using a crowd-sourced platform or not.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/smd-1} \caption{Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of a couple studies, large effect studies are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95\% CI that did not overlap 0; however, in these cases the larger pattern of results was not consistent between original and first replication.  }\label{fig:smd}
\end{figure}

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study (again following \citeproc{ref-boyce2023}{Boyce et al., 2023}).
Though we did not conduct a systematic review of the literature to identify all possible replications, we also consider the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature).

@@

\caption{\label{tab:tab-porig}$P$-original values between different sets of experiments. The primary analysis is between the original result and the meta-analytic aggregation of all replications. All $p$-original values assume an imputed heterogeneity value of $\tau$=.21.}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{p-original comparing between} \\
\cmidrule(l{3pt}r{3pt}){2-6}
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Original and} & \multicolumn{1}{c}{Rescue and} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-6}
Paper & All reps & 1st Rep and Rescue & Rescue & Non-rescue & Other reps\\
\midrule
Birch \& Bloom 2007 & 0.192 & 0.189 & 0.194 & 0.191 & 0.989\\
Child et al. 2018 & 0.669 & 0.669 & 0.641 & 0.858 & 0.778\\
Chou et al. 2016 & 0.055 & 0.055 & 0.101 & 0.005 & 0.231\\
Craig \& Richeson 2014 & 0.145 & 0.145 & 0.001 & 0.302 & 0.020\\
Gong et al. 2019 & 0.262 & 0.262 & 0.057 & 0.511 & 0.228\\
Haimovitz \& Dweck 2016 & 0.068 & 0.018 & 0.018 & 0.180 & 0.867\\
Hopkins et al. 2016 & 0.290 & 0.290 & 0.654 & 0.004 & 0.015\\
Jara-Ettinger et al. 2022 & 0.014 & 0.014 & 0.000 & 0.006 & 0.000\\
Krauss \& Wang 2003 & 0.271 & 0.271 & 0.519 & 0.139 & 0.311\\
Ngo et al. 2019 & 0.106 & 0.000 & 0.000 & 0.385 & 0.257\\
Paxton et al. 2012 & 0.011 & 0.011 & 0.005 & 0.023 & 0.649\\
Payne et al. 2008 & 0.023 & 0.071 & 0.031 & 0.079 & 0.895\\
Porter et al. 2016 & 0.370 & 0.370 & 0.905 & 0.002 & 0.003\\
Schechtman et al. 2010 & 0.712 & 0.712 & 0.654 & 0.770 & 0.877\\
Tarampi et al. 2016 & 0.000 & 0.000 & 0.000 & 0.000 & 0.145\\
Todd et al. 2016 & 0.062 & 0.100 & 0.124 & 0.058 & 0.778\\
Yeshurun \& Levy 2003 & 0.614 & 0.007 & 0.017 & 0.943 & 0.474\\
\bottomrule
\end{tabular}
\end{table}

Scientists' intuitions about whether a replication is successful and whether an effect provides support for a hypothesis (including heuristic cutoffs like p\textless.05) do not always align with measures of statistical consistency (\citeproc{ref-patil2016}{Patil et al., 2016}).
For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population.
On the other hand, one study may find a statistically significant, but imprecisely estimated effect in a small sample, and a second study may find a near-zero (null) effect, but the effect estimates from the two studies may be statistically compatible, despite one supporting a hypothesized difference and the other not.

As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
The value of \(p\)-original represents how likely an effect size equal to or more extreme than the original effect size is to occur if it comes from a distribution of population effect sizes defined by the mean and standard error of the replications (combined meta-analytically) with a between-population standard deviation of \(\tau=.21\) (imputed from \citeproc{ref-olsson-collentine2020}{Olsson-Collentine et al., 2020}).
That is, \(p\)-original is a measure of the statistical consistency of the original result with the totality of the replications, assuming a given level of heterogeneity of effects between studies.
Smaller \(p\)-original values indicate more inconsistency.

@@

As secondary measures, we calculated the \(p\)-original values between a) the original and just the first replication and rescue (i.e.~excluding additional replications), b) the original and the rescue, c) the original and non-rescue replications, and d) the rescue and other replications.
For a) the original versus the first replication and the rescue, the median value of \(p\)-original was 0.1 {[}IQR: 0.01 - 0.27{]}, and 35\% of the \(p\)-original values were less than .05.
For b) the original versus the rescue, the median value of \(p\)-original was 0.06 {[}IQR: 0.01 - 0.52{]}, and 47\% of the \(p\)-original values were less than .05.
For c) the original versus all the non-rescue replications, the median value of \(p\)-original was 0.14 {[}IQR: 0.01 - 0.38{]}, and 35\% of the \(p\)-original values were less than .05.
For d) the rescue versus the other replications, the median value of \(p\)-original was 0.31 {[}IQR: 0.14 - 0.78{]}, and 24\% of the \(p\)-original values were less than .05.

@@

All of the correlations are presented in Table \ref{tab:cor-table}.
The number of rescues is small, and many of these predictors are correlated, so we caution against over interpretation.

The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of the pre-registered correlates meet the conventional significance threshold; the two correlates based on ratios that reflect a relatively smaller replication sample size are marginally significant.

Small replication samples relative to original and rescue could be due to both a) powering a replication according to a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a smaller than intended sample.
Since relative sizes of the studies may play a role in replication success and how probative replications are, we show the sample sizes in Table \ref{tab:tab-size}.

@@

We have focused here on a narrow type of replication and rescue: that done by an early career researcher with the goal of establishing whether this is a result \emph{they} can build on. We acknowledge that this type of replication exists as part of a wider ecosystem of replications that have different properties and are done for different reasons.

>>>>

\newcommand\blfootnote[1]{%
	\begingroup
	\renewcommand\thefootnote{}\footnote{#1}%
	\addtocounter{footnote}{-1}%
	\endgroup
}

\begin{document}
\maketitle
\blfootnote{* Corresponding author. Email: \href{mailto:vboyce@stanford.edu}{\nolinkurl{vboyce@stanford.edu}}}
%%%%%%%%%% START AFTER TITLE PARTIAL %%%%%%%%%%%%%

@@

Imagine you are a graduate student and you run a replication of a study that you are interested in building upon in your research.
The replication fails.
Perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely contains 0, with a \(p\)-value of .3.
Or perhaps the interaction is numerically in the wrong direction and the main effects look different.
Whatever the details of the replication failure, you are left with a question: Should you try again and run a re-replication, or should you give up and pick a different study to build upon?

In psychology, large-scale replication projects have found that around half of studies successfully replicate.
Across 100 studies with positive results, the reproducibility project in psychology (RP:P) replicated 36\%-47\% depending on the metric for replication success (\citeproc{ref-opensciencecollaboration}{Open Science Collaboration, 2015}).
With large multi-site samples, Many Labs 1 replicated 11/13 effects (84.6\%), Many Labs 2 replicated 14/28 effects (50\%), and Many Labs 3 replicated 3/10 effects (30\%) (\citeproc{ref-ebersole2016}{Ebersole et al., 2016}; \citeproc{ref-klein2014}{Klein et al., 2014}; \citeproc{ref-klein2018}{Klein et al., 2018}).
Camerer et al. (\citeproc{ref-camerer2018}{2018}) examined 21 behavioral social science studies and successfully replicated 12-14 of them (57\%-67\%) depending on the metric used.
Boyce et al. (\citeproc{ref-boyce2023}{2023}) reported an average subjective replication score of 49\% for 176 replications primary in psychology.

Psychology is not the only discipline where many studies do not replicate.
Large-scale replication projects in other disciplines have found 39/97 of studies with positive effects (40\%) replicating in cancer biology (\citeproc{ref-errington2021}{Errington et al., 2021}), 11/18 studies (61\%) replicating in economics (\citeproc{ref-camerer2016}{Camerer et al., 2016}), and 31/40 studies (78\%) replicating in experimental philosophy (\citeproc{ref-cova2021}{Cova et al., 2021}).
Across these disciplines, and especially in psychology, large-scale replications indicate that the probability of scientists failing to replicate published findings is substantial.

Replications can fail for many reasons. Some reasons for replication failure have to do with the original study, while others have to do with how the replication was conducted.
One potential reason is that the original result is fragile in some way.
It could be a statistically unlikely result that was achieved through chance or \(p\)-hacking, or it could be sensitive to the exact conditions and time it was run under (``hidden moderators'').
Another possibility is that the original reported effect size is inflated due to some combination of heterogeneity, \(p\)-hacking, low power, and publication bias.
If effect sizes are inflated, replication studies that select a sample size with appropriate statistical power for the reported effect size will have too small a sample to reliably detect the (true) smaller effect.
Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Other reasons for replication failure are primarily due to the design of the replication study.
A replication could observe different results from the original due to methodological differences.
These differences include intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that the replicators did not expect to be critical.
A difference in the details of recruitment, such as changing from recruiting participants in person to online, might mean that particular details of a study implementation are no longer appropriate, or the data quality controls are inadequate.
If there are differences in time, place, or subject population between the two studies, researchers may need to change the materials, instructions, or procedure to adapt the experimental paradigm to the new context.
Such adaptation creates more challenges in interpretation: too many changes could cause differing results, but so could too few changes.
In sum, when replications fail, we generally do not know why they failed, although we may speculate.

For the goals of large-scale replication projects, it does not matter why any particular study failed to replicate, because the aim is to estimate the proportion of effects that replicate in a literature.
In contrast, when individual scientists attempt to replicate an effect with the goal of building on it, they care why an individual replication failed.
In an uncertain literature, scientists might not want to commit resources to studying extensions of an effect without first checking that they can measure that effect themselves.
For those individual scientists, it is much more important to know whether the replication failure is due to a fixable issue with their replication attempt or an intrinsic feature of the original result.

While it is possible to speculate after the fact about potential causes of a single replication failure, in most circumstances these causes are hard to discern (absent clear evidence of errors).
Re-replication can serve as a way to test explanations for replication failure.
Re-replications can make modifications to the first replication and give another chance for researchers to measure an effect of interest.
But what are the chances that a re-replication succeeds?

Relatively few studies have estimated re-replication success.
One estimate comes from Ebersole et al. (\citeproc{ref-ebersole2020}{2020}).
After the original RP:P study reported of a replication rate around 40\% (\citeproc{ref-opensciencecollaboration}{Open Science Collaboration, 2015}), Gilbert et al. (\citeproc{ref-gilbert2016}{2016}) raised concerns that methodological differences might explain the low replication rate.
In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication.
Thus, Ebersole et al. (\citeproc{ref-ebersole2020}{2020}) re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts.
This study thus tested whether larger sample sizes and/or ``better'' methodologies would ``rescue'' the failed RP:P replications.
The result was that 2 of the 10 revised protocols found a significant result (but not the 1/11 that originally was successful).
Even when a significant effect was recovered, the effect size was much diminished (\citeproc{ref-ebersole2020}{Ebersole et al., 2020}).
The re-replication success rate was thus 2/9 (22\%) for studies with non-significant first replications under favorable, high-resource circumstances.

While Ebersole et al. (\citeproc{ref-ebersole2020}{2020}) looked at re-replication success rate in a high-resource setting with access to large, multi-site samples and expert input, most early-career scientists do not have comparable resources to devote to re-replication attempts.
On the other hand, early-career researchers may also be prone to a different distribution of causes of initial replication failure.
Thus, it is an open question what the re-replication rate is for failed replications by early-career researchers.
Further, estimating this rate could inform researchers' decisions on what to do next when a replication fails.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate-level class on experimental methods in psychology were given the option to ``rescue'' an experiment that a student in a prior year had previously failed to replicate as their class project (first replications reported in \citeproc{ref-boyce2023}{Boyce et al., 2023}).
The focus of the class was best practices in experimental design; thus, students looked at both the original paper and the failed replication to examine potential issues in the replication that they could fix in their rescue project.
In the present paper, we report the results of 17 rescue projects.
We find that a minority of projects (5/17, 29\%) successfully recovered the original effect.

@@

PSYCH 251 is a graduate-level experimental methods class taught by MCF.
In previous years, students conducted replication projects (\citeproc{ref-boyce2023}{Boyce et al., 2023}; \citeproc{ref-hawkins2018}{Hawkins et al., 2018}).
In Fall 2023, students were offered the option to complete a ``rescue'' project where they re-replicated one of the unsuccessful replications from a previous year.
Students could also opt to do a normal replication instead.
If they opted for a rescue, they were given the option to participate in a collaborative meta-science project and earn authorship on the current paper based on 1) completing a rescue project, 2) reviewing another student's rescue project for reproducibility, and 3) reviewing and approving the final manuscript.
We report on the result of the 17 rescue projects by students who completed data collection and opted to be a part of the paper.

A spreadsheet of projects, individual project write-ups (both first replications and rescues), individual project data and analyses for rescue projects, and the analytic code for this paper are all available at \url{https://osf.io/cyk5w}.

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is pre-registered at \url{https://osf.io/5qz7v}, individual pre-registrations are linked from \url{https://osf.io/r2d6q}.
We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance and the variance of the comparison (replication or rescue) study (\citeproc{ref-patil2016}{Patil et al., 2016}).
Thus we cannot show a single prediction interval for the original study, but would have to show a prediction interval between each pair of studies.
We thought this analysis would not offer additional clarity and hence we did not perform it.

@@

Students were required to think of reasons why the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions was given to students (\url{https://osf.io/dfwk2}).
In general, we encouraged students to add manipulation checks as appropriate and to adapt their materials to online studies, which may benefit from attention checks and/or shorter duration.
For instance, the rescue of Paxton et al. (\citeproc{ref-paxton2012}{2012}) switched from using the Cognitive Reflection Test (CRT) which many online participants are over-familiar with to using the newer CRT-2, which is used less but attempts to measure the same construct (\citeproc{ref-thomson2016}{Thomson \& Oppenheimer, 2016}).
The rescue of Jara-Ettinger et al. (\citeproc{ref-jara-ettinger2022}{2022}) discovered that the replication had accidentally used an illustrated version of the stimuli rather than the photographic stimuli used in the original and used the photographic stimuli instead.
Tarampi et al. (\citeproc{ref-tarampi2016}{2016}) had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map.
The replication ported this experiment online and had participants use a dropdown menu to select left or right for each turn.
The rescue removed the unwieldy dropdown menus and opted to have participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once experimental designs and analytic plans were approved by class TAs (VB and BP), students pre-registered and ran their samples.

With one exception, samples were collected on Prolific (the rescue of \citeproc{ref-yeshurun2003}{Yeshurun \& Levy, 2003} ran in person on the Stanford student subject pool).
We tried to power studies adequately (with a target sample size of 2.5x the original study following \citeproc{ref-simonsohn2015}{Simonsohn, 2015}), but due to cost constraints, not all studies were powered at this level.
The rescues had on average 1.48 times the original sample post-exclusion (median: 1.07, IQR: 0.94 - 2.4, minimum: 0.48, maximum: 2.96, see Table \ref{tab:tab-size} for all sample sizes).
Across the 16 Prolific studies, we spent \$5471, for an average of \$342 per project.

@@

As a complement to the subjective rating of overall success, we statistically compared one key measure of interest for each study, following Boyce et al. (\citeproc{ref-boyce2023}{2023}).
In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, first replication, and rescue.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data to obtain comparable values.

@@

We recorded the same set of potential correlates that were used in Boyce et al. (\citeproc{ref-boyce2023}{2023}) for original, first replication, and rescue (these were already rated for original and first replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it used a within- or between-subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

For the original study and each replication, we recorded the number of participants after exclusions.
For studies where extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted using an online crowdsourcing platform or not.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/smd-1} \caption{Standardized effect sizes of original studies, first replications, rescues, and additional replications if available. Due to the large effect size of some studies, studies with large effects are shown in a separate panel. In a few cases, the first replication's key effect was in the same direction as the original with a 95\% CI that did not contain 0. However, in these cases the larger pattern of results was not consistent between original and first replication.}\label{fig:smd}
\end{figure}

As a complement to subjective replication ratings, we also statistically compared the effect sizes of the rescue, first replication, and original study on one preregistered key measure per study (again following \citeproc{ref-boyce2023}{Boyce et al., 2023}).
Though we did not conduct a systematic review of the literature to identify all possible replications, we also considered the effect sizes obtained in additional replications when we were aware of them (either from other class projects, or external replications in the literature).

@@

\caption{\label{tab:tab-porig}$P$-original values between different sets of experiments. The primary analysis is between the original result and the meta-analytic aggregation of all replications. All $p$-original values assume an imputed heterogeneity value of $\tau$=.21.}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{p-original comparing between} \\
\cmidrule(l{3pt}r{3pt}){2-6}
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Original and} & \multicolumn{1}{c}{Rescue and} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-6}
Paper & All reps & 1st Rep and Rescue & Rescue & Non-rescue & Other reps\\
\midrule
Birch \& Bloom 2007 & 0.192 & 0.189 & 0.194 & 0.191 & 0.989\\
Child et al. 2018 & 0.669 & 0.669 & 0.641 & 0.858 & 0.778\\
Chou et al. 2016 & 0.055 & 0.055 & 0.101 & 0.005 & 0.231\\
Craig \& Richeson 2014 & 0.145 & 0.145 & 0.001 & 0.302 & 0.020\\
Gong et al. 2019 & 0.262 & 0.262 & 0.057 & 0.511 & 0.228\\
Haimovitz \& Dweck 2016 & 0.068 & 0.018 & 0.018 & 0.180 & 0.867\\
Hopkins et al. 2016 & 0.290 & 0.290 & 0.654 & 0.004 & 0.015\\
Jara-Ettinger et al. 2022 & 0.009 & 0.009 & 0.000 & 0.004 & 0.000\\
Krauss \& Wang 2003 & 0.271 & 0.271 & 0.519 & 0.139 & 0.311\\
Ngo et al. 2019 & 0.106 & 0.000 & 0.000 & 0.385 & 0.257\\
Paxton et al. 2012 & 0.011 & 0.011 & 0.005 & 0.023 & 0.649\\
Payne et al. 2008 & 0.023 & 0.071 & 0.031 & 0.079 & 0.895\\
Porter et al. 2016 & 0.370 & 0.370 & 0.905 & 0.002 & 0.003\\
Schechtman et al. 2010 & 0.712 & 0.712 & 0.654 & 0.770 & 0.877\\
Tarampi et al. 2016 & 0.000 & 0.000 & 0.000 & 0.000 & 0.145\\
Todd et al. 2016 & 0.062 & 0.100 & 0.124 & 0.058 & 0.778\\
Yeshurun \& Levy 2003 & 0.614 & 0.007 & 0.017 & 0.943 & 0.474\\
\bottomrule
\end{tabular}
\end{table}

Scientists' intuitions about whether a replication is successful and whether an effect provides support for a hypothesis (including heuristic cutoffs like p\textless.05) do not always align with measures of statistical consistency (\citeproc{ref-patil2016}{Patil et al., 2016}).
For instance, two studies may both find that condition 1 results in a significantly higher outcome measure than condition 2, but the effect magnitudes may be sufficiently different that they are statistically unlikely to have come from the same population.
On the other hand, one study may find a statistically significant, but imprecisely estimated, effect in a small sample, and a second study may find a near-zero (null) effect. The effect estimates from the two studies might be statistically compatible, despite one supporting a hypothesized difference and the other not.

As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
The value of \(p\)-original represents how likely an effect size equal to or more extreme than the original effect size is to occur if it comes from a distribution of population effect sizes defined by the mean and standard error of the replications (combined meta-analytically) with a between-population standard deviation of \(\tau=.21\) (imputed from \citeproc{ref-olsson-collentine2020}{Olsson-Collentine et al., 2020}).
That is, \(p\)-original measures the statistical consistency of the original result with the totality of the replications, assuming a given level of heterogeneity of effects between studies.
Smaller \(p\)-original values indicate more inconsistency.

@@

As secondary measures, we calculated the \(p\)-original values between a) the original and the first replication and rescue (i.e.~excluding additional replications), b) the original and the rescue, c) the original and non-rescue replications, and d) the rescue and other replications.
For a) the original versus the first replication and the rescue, the median value of \(p\)-original was 0.1 {[}IQR: 0.01 - 0.27{]}, and 35\% of the \(p\)-original values were less than .05.
For b) the original versus the rescue, the median value of \(p\)-original was 0.06 {[}IQR: 0.01 - 0.52{]}, and 47\% of the \(p\)-original values were less than .05.
For c) the original versus all the non-rescue replications, the median value of \(p\)-original was 0.14 {[}IQR: 0 - 0.38{]}, and 35\% of the \(p\)-original values were less than .05.
For d) the rescue versus the other replications, the median value of \(p\)-original was 0.31 {[}IQR: 0.14 - 0.78{]}, and 24\% of the \(p\)-original values were less than .05.

@@

All of the correlations are presented in Table \ref{tab:cor-table}.
The number of rescues is small, and many of these predictors are correlated, so we caution against over-interpretation.

The strongest correlates of rescue success were open materials, a small sample size in the first replication, a small sample size in the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of the pre-registered correlates meet the conventional significance threshold; the two correlates based on ratios that reflect a relatively smaller replication sample size are marginally significant.

Small replication samples relative to original and rescue could be due to both a) powering a replication according to a reported large effect size or b) difficulties with recruitment or high exclusion rates leading to a sample size smaller than intended.
Since relative sizes of the studies may play a role in replication success and how probative replications are, we show the sample sizes in Table \ref{tab:tab-size}.

@@

We have focused here on a narrow type of replication and rescue: that done by an early-career researcher with the goal of establishing whether \emph{they} can build on a result. We acknowledge that this type of replication exists as part of a wider ecosystem of replications that have different properties and are done for different reasons.
